---
title: "Convergence checks"
date: "`r format(Sys.time(), '%d %B %Y, %H:%M')`"
output:
  html_notebook:
    code_folding: hide
    toc: yes
    toc_float: yes
---
<style type="text/css"> <!-- .table { width: auto } ---> </style>

```{r options, include=FALSE}
knitr::opts_chunk$set(
  # results = "hide",
  message = FALSE,
  warning = FALSE,
  package.startup.message = FALSE
  )

options(
  knitr.kable.NA = "",
  width = 100
)
```

```{r}
library(jagsUI)

wp_model <- 2  # set Model to estimate -- M1 or M2

file_wikipedia <- "../01-data-sources/02-wikipedia/wikipedia-data/01-wp-data-json.zip"
file_wikipedia_date <- substr(file.info(file_wikipedia)$mtime, 1, 10)

file_estimation <- sprintf("estimation-model/02-result-m%d.RData", wp_model)
file_estimation_date <- substr(file.info(file_estimation)$mtime, 1, 10)
```

__Model `r wp_model`__ — M1 ideology tags only, M2 ideology and position tags

+ Wikipedia data download -- `r file_wikipedia_date`
+ Model estimation -- `r file_estimation_date`
+ [ date file modified ]


## Convergence diagnosis

```{r}
load(sprintf("estimation-model/02-result-m%d.RData", wp_model))
```


```{r}
params <- c("alpha", "beta", "o", "x")

if (wp_model == 2) {
  
  params <- c(c("gamma", "tau"), params)
  
}

pl_params <- substr(params, 1, 1)
n_params <- length(params)
```

```{r}
get_df <- function(var) {

  quantile_select <- c(0.5, 0.75, 0.9, 0.95, 0.99)

  if (var == "n.eff") {
    
    quantile_select <- abs(rev(quantile_select) - 1)
    
  }

  dt <- data.frame(
    min = round(sapply(jags_out[[var]][1:n_params], min), 3),
    mean = round(sapply(jags_out[[var]][1:n_params], mean), 3),
    max = round(sapply(jags_out[[var]][1:n_params], max), 3),
    quant = t(
      round(sapply(jags_out[[var]][1:n_params], quantile, quantile_select), 3)
      ),
    Npar = sapply(jags_out[[var]][1:n_params], length)
  )
  
  names(dt) <- gsub("quant\\.", "", names(dt))
  names(dt) <- gsub("\\.$", "%", names(dt))
  rownames(dt) <- params
  
  return(dt)
}
```

Scale reduction factors $\hat{R}$ --- close to one and all values < 1.1.

```{r}
get_df("Rhat")
```

Effective samples sizes $N_{eff}$ --- all values > 100

```{r}
get_df("n.eff")
```


## SD / Mean checks

Confirming $\bar{x} = 0$

```{r}
round(summary(apply(jags_out$sims.list$x, 1, mean)), 3)
```

Confirming $\text{SD}(x) = 1$

```{r}
round(summary(apply(jags_out$sims.list$x, 1, sd)), 3)
```


## Distribution parties

```{r}
plot(density(jags_out$mean$x), main = "")
```


## Traceplots

Inspect traceplots for particular parameters

__Plots only presented for critical values__

-----

```{r}
check_condition <- function(var_1, operator, var_2) {
  
   operator(var_1, var_2)
  
}

get_plots <- function(var, operator, value) {
  
  pars <- list()
  
  for (l in pl_params) {
    
    pars[[l]] <- sapply(
      which(check_condition(jags_out[[var]][[l]], operator, value)), 
      function(.x) paste0(l, "[", .x, "]")
    )
    
    if (! is.character(pars[[l]])) {
      
      pars[[l]] <- NULL 
      
    }
    
  }
  
  pars <- unlist(pars)
  
  if (! is.null(pars)) {
    
    lapply(pars, function(.x) traceplot(jags_out, parameters = .x))
    
  }
}
```

### R-hat

Parameters with $\hat{R} > 1.1$

```{r}
get_plots("Rhat", `>`, 1.1)
```

### N-eff

Parameters with $N_{eff} < 100$

```{r}
get_plots("n.eff", `<`, 100)
```

